---
title: "Index"
author: "Irene"
date: "2024-11-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)

library(tidyverse)
library(here)
install.packages('gganimate')
library(gganimate)
install.packages("gifski")

```

## Project Description
This project involves the visualisation of Sheffield's weather data. 

The data is retrieved from: https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/sheffielddata.txt

The URL for these pages is:https:
//ijames8.github.io/sheffield_weather_data

The repository for these pages is: 
https://github.com/IJames8/sheffield_weather_data.git



## Import the Data

```{r import}
loc = "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/sheffielddata.txt"

# split values by 'tabs' with \t. Tell R to not split the values by commas, but to split it up by tab.

# this approach does not work. 

# df <- read.csv(loc, sep = "\t", skip = 7)

# import as text, instead of csv, and clean.

dat <- readLines(loc)

# change the word "provisional" to nothing, replace every instance. 
dat <- gsub("Provisional", "", dat)
dat <- gsub("\\*", "", dat)
dat <- gsub("\\#", "", dat)

# save new file
writeLines(dat,here("data", "cleaned.txt" ))

# import data file

df <- read.table("data/cleaned.txt", skip = 7, fill = TRUE)

```


## Wrangle the Data
```{r wrangle}
cols <- c("year", "month", "max_t", "min_t", "af", "rain", "sun")

colnames(df) <- cols

```



## Visualise the Data

```{r plot}

# I need max temperature to be a number not character. 

df$max_t <- as.numeric(df$max_t)

# create plot, change transparency with alpha. 
# canvas

p <- ggplot(df, aes( x = month, y = max_t, group = year, colour = year))

p + geom_line(alpha = 0.1) +
  geom_line(data = df %>% filter(year ==2024), colour = "red", linewidth = 2)



```


##Animated plot

We are going to use gganimate to create an animated plot. Consider if the animation is actually going to be helpful to the data visualisation. 

```{r animate}
# create an average temp for each year

df2 <- df %>%
  group_by(year) %>%
  summarise(avg_t = mean(max_t))

# create an animated plot

p2 <- ggplot(df2, aes( x = year, y = avg_t), frame = year)

# Cumulative keeps it on the screen
anim <- p2 + geom_point() +
  transition_manual(year, cumlulative = TRUE)+
  ggtitle("The Year is: {frame + 1882}")

anim

# don't think I downloaded gganimate properly. 

# export as gif, then save the plot to be embedded into github pages. 
  
```




